### 2015-03-03 FirstDraftMaterials/Outline.tex
adding outline

\documentclass{article}

\usepackage{amsmath}
\usepackage{mathtools}
\setlength\parindent{0pt}

\begin{document}

\section{Introduction}
Kelsey
\section{LWR Model}
Kelsey
\section{PW Model}

\subsection{Theory}
Kelsey
\subsection{Hugoniot Loci and Integral Curves}
Brisa
\section{AR Model}

\subsection{Theory}
Kelsey
\subsection{Hugoniot Loci and Integral Curves}
Brisa

\section{Examples}
Brisa

\end{document}